---
title: "COD_20231129_MGK_IBS7048_week14_data_visualization_advanced"
author: "Minsik Kim"
date: "2023-11-29"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- 
         ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                "/Volumes/MacMini Drive/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/", 
                "/Users/minsikkim/Dropbox (Personal)/Inha/5_Lectures/Advanced biostatistics/scripts/BTE3207_Advanced_Biostatistics/")

path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("tidyverse", "pacman", "yaml", "knitr", "rmarkdown", "rmdformats")


YAML_header <-
        '---
title: "COD_20231129_MGK_IBS7048_week14_data_visualization_advanced"
author: "Minsik Kim"
date: "2032.11.29"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230920"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```



# Roading ggplot

from previous lecture.. we generated a phyloseq object. it can be saved to hard disk using the below command.

```{r}

library(ggplot2)
library(tidyverse)

```


# Advanced ggplot2

```{r}

head(mpg)

```

Generating plot

```{r}
plot(mpg$cty, mpg$hwy)




mpg %>% 
        ggplot(aes(x = cty, y = hwy))

```



```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy)) +
        geom_point()

```



```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)")

```

```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer, shape = as.factor(cyl))) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)")

```

## + theme()

```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.position = "left")

```

```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(axis.title.x = element_text(family = "serif", size = 15))

```
       
## theme() and element_markdown()

```{r}
install.packages("ggtext")
library(ggtext)

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(axis.title.y = element_markdown(family = "serif", size = 15))


```


*Basic markdown syntax is at https://www.markdownguide.org/basic-syntax/*

### Italic

```{r}


mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("*Highway mileage (MPG)*") +
        theme(axis.title.y = element_markdown(family = "serif", size = 15))

```

### Bold

```{r}


mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("**Highway mileage (MPG)**") +
        theme(axis.title.y = element_markdown(family = "serif", size = 15))

```

### Italic

```{r}


mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("*Highway mileage (MPG)*") +
        theme(axis.title.y = element_markdown(family = "serif", size = 20))

```

### superscript and subscript

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("High<sup>way</sup>
             mileage 
             <sub>(MPG)</sub>") +
        theme(axis.title.y = element_markdown(family = "serif", size = 25))

```

### Comprehensive

```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("*Hi***gh**<sub>way</sub> mile<sub>age</sub> ***(MPG)***") +
        theme(axis.title.y = element_markdown(family = "serif", size = 30))

```


## + guides()

```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        guides(col=guide_legend(nrow=3))

```



```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        guides(col=guide_legend(ncol=3, title.position = "bottom"))

```


```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        guides(col=guide_legend(ncol=3, title.position = "bottom",
                                override.aes = list(size = 4)))


```



```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        guides(col=guide_legend(ncol=3, title.position = "top", 
                                title.hjust = 0.5))
```

```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3,
                                title.position = "top",
                                title.hjust = 0.5,
                                keywidth = 0.2,
                                keyheight = 0.2,
                                title = "*Manufacturer*<sup>company</sup>"))
```


## theme_...


### classic 

The order matters (theme and theme_classic)

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3, title.position = "bottom", title.hjust = 1, keywidth = 0.2, keyheight = 0.2, title = "*Manufacturer*<sup>company</sup>")) +
        theme_classic() 
```

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() + 
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3, title.position = "bottom", title.hjust = 1, keywidth = 0.2, keyheight = 0.2, title = "*Manufacturer*<sup>company</sup>"))
```
### bw



```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_bw() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3, title.position = "bottom", title.hjust = 1, keywidth = 0.2, keyheight = 0.2, title = "*Manufacturer*<sup>company</sup>"))
```
### dark

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_dark() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3, title.position = "bottom", title.hjust = 1, keywidth = 0.2, keyheight = 0.2, title = "*Manufacturer*<sup>company</sup>"))
```
### minimal

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_minimal() + 
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3, title.position = "bottom", title.hjust = 1, keywidth = 0.2, keyheight = 0.2, title = "*Manufacturer*<sup>company</sup>"))
```

## geom_smooth

### lm, one line

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3, title.position = "bottom", title.hjust = 1, keywidth = 0.2, keyheight = 0.2, title = "*Manufacturer*<sup>company</sup>")) +
        geom_smooth(method = "lm", aes(x = cty, y = hwy, col = NA))


```
### lm multiple line

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3, title.position = "bottom", title.hjust = 1, keywidth = 0.2, keyheight = 0.2, title = "*Manufacturer*<sup>company</sup>")) +
        geom_smooth(method = "lm", aes(x = cty, y = hwy, col = manufacturer))


```

### `se`, Confidence intervals



```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3, title.position = "bottom", title.hjust = 1, keywidth = 0.2, keyheight = 0.2, title = "*Manufacturer*<sup>company</sup>")) +
        geom_smooth(method = "lm", aes(x = cty, y = hwy, col = manufacturer),
                    se = F)


```

### gam, curvy line

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(ncol=3, title.position = "bottom", title.hjust = 1, keywidth = 0.2, keyheight = 0.2, title = "*Manufacturer*<sup>company</sup>")) +
        geom_smooth(method = "gam", aes(x = cty, y = hwy, col = NA))



```


## Colors

### manual colors


```{r}


mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c("#e5f5f9", "#99d8c9", "#2ca25f", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey"))



```

### Changing labels

```{r}


mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c("red", "blue", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey"), 
                           labels = c("Audi", "Chevy", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey", "grey"))



```

### Automated colors1

```{r}


mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_viridis_d()



```

### Automated colors2

```{r}


mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_brewer()



```

### Automated colors3

```{r}


mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_brewer(type = "qual", palette = 6)



```

### Automated colors4

```{r}
mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_brewer(type = "div", palette = 2)


```

Note https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3 for more colors, that they are distinguishable.

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999'))

```

## facetting


## facet_wrap

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_wrap(~manufacturer)

```
```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_wrap(~cyl, nrow = 1)

```


```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_wrap(~cyl, nrow = 1, scales = "free_y")

```

### facet_grid

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_grid(year~cyl)

```
## Title

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_grid(year~cyl) +
        ggtitle("Correlation between city mpg and high mpg")

```

## labs

```{r}

mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_grid(year~cyl) +
        labs(tag = "A")

```

## ggarrange


```{r}
#install.packages("ggpubr")
library(ggpubr)

ggarrange(
        mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        labs(tag = "A")
        
        ,
        
        
        
        mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown(),
              axis.text.x = element_blank()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_grid(year~cyl) +
        labs(tag = "B")
        
        
)



```
### common.legend

```{r}
ggarrange(
        mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        labs(tag = "A")
        
        ,
        
        
        
        mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown(),
              axis.text.x = element_blank()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_grid(year~cyl) +
        labs(tag = "B")
        
        ,
        
        common.legend = T
        
)



```

### widths = c()




```{r}

plot1 <-
        mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        labs(tag = "A")
        

plot2 <- mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown(),
              axis.text.x = element_blank()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_grid(year~cyl) +
        labs(tag = "B")





ggarrange(plot1, plot2, common.legend = T, widths = c(1,3))
        
                


```

## Saving plot file

saving plot as an object

```{r}
plot_ab <- ggarrange(
        mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        labs(tag = "A")
        
        ,
        
        
        
        mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown(),
              axis.text.x = element_blank()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_grid(year~cyl) +
        labs(tag = "B")
        
        ,
        
        common.legend = T
        
        
        
        ,
        
        widths = c(1,3)
        
                
        
)


plot_ab

```

### Final assignment


```{r}


plot_abc <- ggarrange(#QUESTION 1
        mpg %>%
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() + #QUESTION 2
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) + ##QUESTION 3
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) + #QUESTION 4
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        labs(tag = "A")
        
        ,
        
        
        
        mpg %>% 
        ggplot(aes(x = cty, y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_point() +
        xlab("City mileage (MPG)") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown(),
              axis.text.x = element_blank()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        facet_grid(year~cyl) + #QUESTION 5
        labs(tag = "B")
        
        ,
        
        
        
        mpg %>% 
        ggplot(aes(x = as.factor(cyl), y = hwy, col = manufacturer)) +
        theme_classic() +
        geom_boxplot() +
        xlab("Number of cylinders") +
        ylab("Highway mileage (MPG)") +
        theme(legend.title = element_markdown()) +
        guides(col=guide_legend(title = "Manufacturer", override.aes = list(size = 4))) +
        scale_color_manual(values = c('#8dd3c7','#ffffb3','#bebada','#fb8072','#80b1d3','#fdb462','#b3de69','#fccde5','#d9d9d9',
                             '#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628','#f781bf','#999999')) +
        #facet_grid(year~cyl) + #QUESTION 5
        labs(tag = "C")
        
        , 
        
        
        ncol = 3,
        
        
        common.legend = T #QUESTION 6
        
        
        
        ,
        
        widths = c(1,1,2)#QUESTION 7
        
                
        
)


plot_ab

```


### pdf

```{r}
#getwd()
#dir.create("Figures")

pdf(file = "Figures/Figure1.pdf",   # The directory you want to save the file in
    width = 10, # The width of the plot in inches
    height = 10 # The height of the plot in inches

) #fixing multiple page issue

plot_ab


dev.off()


```



### png

```{r}
#dir.create("Figures")

png(file = "Figures/Figure1.png",   # The directory you want to save the file in
    width = 180, # The width of the plot in inches
    height = 170, # The height of the plot in inches
    units = "mm",
    res = 600
) #fixing multiple page issue

plot_ab


dev.off()




```

How to use kbl

# kableExtra package


```{r}

#devtools::install_github("kupietz/kableExtra")
install.packages("kabelExtra")

library(kableExtra)

```

## kbl



```{r}
mpg


mpg %>% 
        head() %>% 
        kbl(format = "html", escape = 0)


```

## kable_styling



```{r}

mpg %>% 
        head() %>% 
        kbl(format = "html", escape = 0) %>%
        kable_styling(full_width = 0, html_font = "sans") 




```

Application

```{r}




audi <- mpg %>% 
        filter(manufacturer == "audi") %>%
        lm(data = ., cty ~ year + cyl + displ) 

chevy <- mpg %>% 
        filter(manufacturer == "chevrolet") %>%
        lm(data = ., cty ~ year + cyl + displ) 

audi %>% summary
```


```{r}

audi$coefficients 
confint(audi)
paste0(audi$coefficients %>%
               round(2),
       ", (",
       confint(audi)[,1] %>%
               round(2),
       ", ",
       confint(audi)[,2] %>%
               round(2),
       ")")



```


```{r}
summary(audi)$coefficients %>% 
        data.frame(check.names = F) %>% 
        .$`Pr(>|t|)` %>%
        round(., 4) %>%
        format(nsmall = 4)
```






```{r}

audi2 <- cbind (
        `Effect (95% CIs)` = 
               paste0(
                       audi$coefficients %>%
                               round(2),
                       ", (",
                       confint(audi)[,1] %>%
                               round(2),
                       ", ",
                       confint(audi)[,2] %>%
                               round(2),
                       ")"),
       `<i>p</i>-value` = summary(audi)$coefficients %>% 
               data.frame(check.names = F) %>% 
                .$`Pr(>|t|)` %>%
                round(4) %>%
               format(nsmall = 4),
       ` ` = ifelse(summary(audi)$coefficients %>% 
               data.frame(check.names = F) %>% 
                .$`Pr(>|t|)` %>%
                       as.numeric() < 0.05,
               "*",
               "")
       )


chevy2 <- cbind (
        `Effect (95% CIs)` = 
               paste0(
                       chevy$coefficients %>%
                               round(2),
                       ", (",
                       confint(chevy)[,1] %>%
                               round(2),
                       ", ",
                       confint(chevy)[,2] %>%
                               round(2),
                       ")"),
       `<i>p</i>-value` = summary(chevy)$coefficients %>% 
               data.frame(check.names = F) %>% 
                .$`Pr(>|t|)` %>%
                round(4) %>%
               format(nsmall = 4),
       ` ` = ifelse(summary(chevy)$coefficients %>% 
               data.frame(check.names = F) %>% 
                .$`Pr(>|t|)`  %>%
                       as.numeric() < 0.05,
               "*",
               "")
       )

data <- cbind(audi2, chevy2) %>% 
        as.data.frame()
        
data

row.names(data) <- c("Intercept",
                     "Year produced<sup>a</sup>",
                     "Number of cylinders",
                     "Displacement")

data 


```

https://blog.hubspot.com/website/how-to-bold-in-html#:~:text=To%20italicize%20the%20text%20in,style%20property%20set%20to%20italic.


```{R}
data %>%
        kbl(format = "html", escape = 0) %>% 
        kable_styling(full_width = 0, html_font = "sans") %>%
        add_header_above(c(" " = 1, "Audi" = 3, "Chevrolet" = 3)) %>% 
        kable_styling(full_width = 0, html_font = "sans") 

```

# Saving kbl file


```{r}


table <- data %>%
        kbl(format = "html", escape = 0) %>% 
        kable_styling(full_width = 0, html_font = "sans") %>%
        add_header_above(c(" " = 1, "Audi" = 3, "Chevrolet" = 3)) %>% 
        add_footnote("Only data from 1999 and 2008 were used.") %>%
        kable_styling(full_width = 0, html_font = "sans")
        
table

save_kable(table, "Figures/Table1.html")


```

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```