---
title: "COD_20241002_MGK_IBS7048_week5_introduction_to_R"
author: "Minsik Kim"
date: "2024-10-02"
output:
        rmdformats::downcute:
        downcute_theme: "chaos"
code_folding: show
fig_width: 6
fig_height: 6
df_print: paged
editor_options: 
        chunk_output_type: inline
markdown: 
        wrap: 72
---
        
```{r warning=FALSE, message=FALSE, echo=FALSE, results='hide', setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)


path_working <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/mnt/4T_samsung/Dropbox/",
               ifelse(sessionInfo()[1]$R.version$platform == "aarch64-apple-darwin20",
                      "/Volumes/macdrive/Dropbox/", 
                      ""))
path_library <- 
        ifelse(sessionInfo()[1]$R.version$platform == "x86_64-pc-linux-gnu",
               "/home/bagel/R_lib/",
               "/Library/Frameworks/R.framework/Resources/library")



str_libraries <- c("tidyverse", "reactable", "yaml", "pacman")




YAML_header <-
        '---
title: "COD_20241002_MGK_IBS7048_week4_introduction_to_R"
author: "Minsik Kim"
date: "2024.10.02"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'

seed <- "20240902"


#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
        .libPaths(path_library)
        
        require(pacman)
        pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))
        
        knitr::opts_knit$set(root.dir = path_working)
        
        str_libraries |> unique() |> sort() -> str_libraries
        pacman::p_load(char = str_libraries)
        
        set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
        cat("Line Zero Environment:\n\n")
        paste("R:", pacman::p_version(), "\n") |> cat()
        cat("Libraries:\n")
        for (str_libraries in str_libraries) {
                paste(
                        "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
                        "\n", sep = ""
                ) |> cat()
        }
        paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
        paste("    Library Path:", path_library, "\n") |> cat()
        paste("    Working Path:", path_working, "\n") |> cat()
        paste("Seed:", seed, "\n\n") |> cat()
        cat("YAML Header:\n")
        cat(YAML_header)
}
FUN.LineZero.Report()


```

# Working with data in R

There are different type of data in R. The smallest data is called `data element`, and when data element is gathered (with orders), that chunk of data is called `vector`.

Here are some examples.

## Data element

```{r}
alpha <- 1 # this is data element

alpha # if we call alpha, computer will tell you what is assaigned as alpha.

```

## Vecetor

To assign vector in R, we use `c()`, with comma `,` as separators. For example, 

```{r}

beta <- c(1, 2) # this is vector

beta

```


```{r}
gamma <- c(alpha, beta) # it is still a vector

gamma

```

Remember vectors are **ordered**.

```{r}

delta <- c(beta, alpha) # it is still a vector

delta

```

*If you call* `beta` *in R, it will find data named as *`beta`. *Then, how do we assign characters (letters, categorical variable)?* **R recognizes character values with **`""`. For instance,

```{r}

delta_2 <- c("beta", "alpha") 

delta_2

```

With quotation marks, the R will recognize inputs as a new character data, not the data we assigned previously.


# Data frame

Then, what should we do with larger data, such as excel spread sheets?

Vectors in R can be chucked together, **when their legnth is the same**.

This is called data frame in R, and it can be assigned using `data.frame()` function.


```{r}

epsilon <- data.frame(gamma, delta) # this is data frame
# note that the length of both gamma and delta is the same.

epsilon

```


## Selecting variables (vectors) in data frame

Data frame has columns, where each column is one variable (vectors)

we use dollar sign to select one variable from data frame

```{r}

epsilon$gamma

epsilon$delta


```

`epsilon$gamma` returns exact the same vector as just `gamma`, which was used for constructing epsilon.

Then, how can we generate larger dataset to analyze something?

## Data frame example

We can make data by adding multiple data elements to vectors, and we can bind those vectors to make one dataframe.


```{r}
subject <- c("Joe", "Trump", "Obama", "George") #assigning multiple data as a vector
height <- c(183, 190, 187, 182)
IsTall <- c("short", "tall", "tall", "short")

example_dataframe <- data.frame(subject, height, IsTall) #binding 3 vectors as one data frame

example_dataframe #loading the example_data frame

```


# Before begin..

*Basics of R*

## Get current pathway of R working environment

```{r}
getwd()
```


## Listing files in current pathway

```{r}
list.files()
```

## Changing directory

Use `tab` for navigating folders!

```{r}
#setwd("dataset/")

getwd()
```

Now, you are in dataset folder!

## Going back to original folder

To go back to the upper path,

```{r}
setwd("..")

getwd()
```

Now you came back to the original working directory.


# Logical values...having logical values.

```{r}
a = 1

a

```

`=` does the same thing as `<-`. 

To test if the thing are same, R uses `==`.


```{r}
a == 1

```

as we inserted `a <- 1` in the previous code chunk, this test results in `TRUE`

```{r}
"a" == 1

```
This test will test whether a character, `"a"`, is the same with a numeric value `1`. As they are not the same, it returns `FALSE`

Here, `TRUE` and `FALSE` results are , and they are one type of **binary variable**. It works for longer vectors or variables as well.

```{r}

c(1, 2, 3, 4, 5) == c(1, 2, 2, 4, 5)

c(1, 2, 3, 4, 5) == 1

```

And it results in a vector of all the logical tests. Using this, we can filter data easily!


# Howe to select values

To select values from vector, we use `[]` (square brackets).

```{r}
a <- c(1, 2, 3)

a[1]

```

`a[1]` will result in the first data element in this vector.

It is slightly different with some data with names.

```{r}

names(a) <- c("first", "second", "third")

str(a)

```
now the vector `a` is named numeric vector.

In this case,

```{r}
a[1] 

```

The results will be the name and the value!

This sometimes causes some problem when calculating data.

In that case, we need to use double brackets`[[]]`

```{r}
a[[1]]
```


By the way, selecting multiple numbers can be done with colons`:`.

```{r}
1:10

5:20
```

As it has output of vector (multiple elements), both below codes will work the same way

```{r}

a[c(1, 2, 3)]

a[1:3]

```

For selecting data in data frames, it works the same way but it separates the rows and columns using comma. Here is one example.

```{r}
dataframe_example <- data.frame(Joe = c(1:1000),
                                Trump = sample(1:1000, 100),
                                Obama = sample(1:1000, 100),
                                George = sample(1:1000, 100)
                                )

head(dataframe_example)
```

This is a data frame, meaning nothing (it just has 100 numbers randomly selected from a numbers between 1 to 1000), besides Biden. It has ordered numbers from 1 to 100.


To select some data, we use numbers or columns again. But we separate inputs with a comma `,`.

Selecting 1st row and 1st cloumn

```{r}

dataframe_example[1,1]



```

Selecting multiple rows in column 1

```{r}

dataframe_example[1:10, 1]



```


Selecting multiple rows and columns

```{r}
dataframe_example[3:5, 1:2]
```


## Loading data

However, we are not going to type all the data whenever they are needed. To save our previous time, we will directly load the data frame from our computer (most commonly `.csv`, we can also import `.xlsx`) using `read.csv()` function.

Here is the example of a SBP (systolic blood pressure) data from https://nhiss.nhis.or.kr/bd/ab/bdabf003cv.do.

```{r}

download.file("https://raw.githubusercontent.com/minsiksudo/BTE3207_Advanced_Biostatistics/refs/heads/main/dataset/sbp_dataset_korea_2013-2014.csv",
              destfile = "sbp_dataset_korea_2013-2014.csv")

list.files()

dataset_sbp <- read.csv(file = "sbp_dataset_korea_2013-2014.csv") 



```

The data is loaded as below. However, as it contains 1M lines of data, calling the data will make this document too long. Here, we can use `head()` or `tail()` functions to see only 10-first or 10-last dataset, respectively.

## head() and tail() functions {.tabset}

### head() 

```{r}

head(dataset_sbp)

```

### tail()

```{r}

tail(dataset_sbp)

```

### full dataset

```{r}

dataset_sbp

```

### {-}

This shows (subsets of) gender, age group, SBP, DBP, FBS, DIS, and BMI data of 1,000,000 subjects. 

For now, we will use dollar sign to select one variable from dataframe. Again, I will use head(function to check starting 10 data at this time)


```{r}

head(dataset_sbp$SBP)

```

Such a long data!

# Basic R functions for statistics

Now, let's use computer to calculate summary statistics of this large dataset.


## Mean calculation

Function for calculating mean is `mean()`

```{r}
mean(dataset_sbp$SBP)
```

This is the mean value of all the SBP data (1,000,000).

## SD calculation

Function for calculating standard deviation is `sd()`

```{r}

sd(dataset_sbp$SBP)

```

This is the SD.

## Median

function for calculating median: median()

Median can be calculated by using `median()`


```{r}

median(dataset_sbp$SBP)

```


## Function for calculating quantile: qunatile()

`quantile()` function has two functions. Without assigning any percentile value of our interest, it will list all the quantiles (0, 25%, 50%, 75% and 100% percentiles) at once.

```{r}

quantile(dataset_sbp$SBP)

```

As the output of `quantile()` shows multiple data element, the output is a **vector**. Kindly, this output also has names as well. To check what the data has output, we can test `str()`. 

```{r}

str(quantile(dataset_sbp$SBP))

```

The str() says it is `named numeric variable`.

## quantile() - continued

However, what should we do if we are curious about values at a specific percentile? In that case we need to give more information to functions.

These information handed over to function are called **arguments**.

As functions in R can have multiple inputs, we separate them with commas (`,`). 

for example, 

10% percentile of the SBP dataset can be calculated using

```{r}

quantile(x = dataset_sbp$SBP, 0.1)

```

# Use question mark to see what function is doing

As thous functions were constructed by multiple , and there are lots of `packages` that can be newly installed to employ new functions in R. Developers listed all the details and instructions when they made new functions, and they can be reviewed by adding `questionmark` to a function.

```{r}
?quantile(data)
```


For some popular functions, details for `arguments` are usullally embeded on pop-ups. Use `tab` button to see what function can do.


# How to install more functions in R

We use `install.packages()` function to install The CRAN (Comprehensive R Archive Network) server.


```{r}
install.packages("tidyverse")
```

However, the package you just installed, is on your computer (somewhere in a folder called libraries), but they are not `loaded` to R program (simply saying, you did not open them). To use the installed packages, you need to use function `library()`

```{r}
library(tidyverse)
```

Now you can use `tidyverse` package!

`tidyverse` package is the one helps you writing code / summarizing results.

When you learn how to install new packages, you can navigate functions in that package using two colons (`::`).

```{r}
tidyverse::tidyverse_conflicts()
```

As we installed a new package, there could be a `conflits` in functions. `tidyverse_conflicts()` shows the list of those conflicts. 

As developers are doing things by their own, and CRAN does not have a system controlling the names of the newly developed functions. That means, a new function from a package that you installed, can be overlapped with other functions from other packages!


```{r}
head(dplyr::filter(dataset_sbp, SEX == 1))

head(stats::filter(dataset_sbp$SEX, rep(1,3)))
```
`dplyr::filter()` will filter out the data based on given condition, `SEX == 1`.

(doing the same thing as `subset()`)

However, `stats::filter()` (which is the basic package comes with R) does different thing. It applies linear filtering to a univariate time series or to each series separately of a multivariate time series.

The best practice is to note all the function names with `::`. But generally, you don't have to do it as it is not that common problem.

# Basic tidyverse

Tidy verse helps you writing code efficiently. But how?


Let's see this example. We want to filter sample, based on some condition. And we have multiple conditions.

```{r}

head(filter(dataset_sbp, SEX == 1 & SBP > 120))

```

This function filtered based on multiple conditions, when `SEX == 1` and `SBP > 120`. But how are we going to do some imputation, and then filter out based on some conditions?


```{r}

head(filter(filter(filter(dataset_sbp, SEX == 1), SBP > 120), FBS > 110))

```

This function filtered based on multiple conditions, when `SEX == 1` and `SBP > 120`. Plus, it has `head` function outside again. 


It can be done with this code and it does the same thing.

```{r}

dataset_sbp %>%
        filter(SEX == 1) %>%
        filter(SBP > 120) %>%
        filter(FBS > 110) %>%
        head()

```


But how are we going to do some imputation, and then filter out based on some conditions?


Let's see this example again.

We can try adding multiple lines of code to do this. Let's say we are interested in the difference between SBP and DBP. And then we want to categorize them with genders. And then, we want to filter out the data based on their quantile.



```{r}

dataset_sbp$Diff_SBP_DBP <- dataset_sbp$SBP - dataset_sbp$DBP

dataset_sbp_male <- filter(dataset_sbp, SEX == 1)
dataset_sbp_female <- filter(dataset_sbp, SEX == 2)

avg_male <- mean(dataset_sbp_male$Diff_SBP_DBP)

avg_female <- mean(dataset_sbp_female$Diff_SBP_DBP)

sd_male <- sd(dataset_sbp_male$Diff_SBP_DBP)

sd_female <- sd(dataset_sbp_female$Diff_SBP_DBP)

data.frame(SEX = c(1, 2),
           aberage_by_group = c(avg_male, avg_female),
           sd_by_group = c(sd_male, sd_female))

```

We did it! However, the codes are quite nasty, and we have generated unnecessary intermediate data frames as well. Isn't there a smarter way?

# Piping

The good news is, `tidyverse::` package has a great feature called *piping*. In basic R, if we do not assign values with `<-`, the computer will just show the result and it won't store the output. 

Piping helps employing that output temprarilly, using `%>%`

```{r}
dataset_sbp %>% head()

```

Selection of piped data in tidyverse can be done with dot `.`.

```{r}
dataset_sbp %>% .$SEX %>% head()

```


The data will be moved the the next function, and will be employed for calculation.

```{r}
dataset_sbp %>% mutate(Diff_SBP_DBP = SBP - DBP) %>% head()
```

See? Here, `mutate()` is a function for calculating new variable in `tidyverse`.

Let's do the same thing with tidyverse.

```{r}

dataset_sbp %>% 
        mutate(Diff_SBP_DBP = SBP - DBP) %>% 
        group_by(SEX) %>%
        summarise(avereage_by_group = mean(Diff_SBP_DBP),
                  sd_by_group = sd(Diff_SBP_DBP))


```

Try to use piping as many as you can. 
And use lots of lines and indents.
It will make your code look tidier.

These application is also possible (I didn't make a bad example as it would take too much of my personal time)

```{r}


dataset_sbp %>% 
        group_by(SEX, DIS) %>%
        summarise(avereage = mean(Diff_SBP_DBP),
                  sd = sd(Diff_SBP_DBP))



```


# Basic visualization of data

We are going to visualize this example dataset (SBP).

## Summary statistics of data

We can use `mean()`, `sd()`, `median()` functions to calculate summary statistics.

```{r}
cat("#Mean of SBP of 1M subject\n",
      mean(dataset_sbp$SBP),
      "\n\n#Standard deviation of SBP of 1M subject\n",
      sd(dataset_sbp$SBP),
      "\n\n#Median of SBP of 1M subject\n",
      median(dataset_sbp$SBP)
      )
```

`cat()` function prints out the character. `\n` changes the line of the console.

Alternatively, `print()` can be used but the output of `print()` function is recognized as data in R. `cat()` just adds lines in the console.


## summary()

Instead, R has a convenient function called `summary()`.

```{r}

summary(dataset_sbp$SBP)

```

Voila! Now we can roughly see how the data looks like. However, it will be more straight forward if we can see the data in a form of figure.

## hist()

`hist()` function creates histogram in R. It as multiple arguments to make more informative histogram as output. For example,

```{r}

hist(dataset_sbp$SBP)

```

## hist() - breaks

`hist()` as argument called `breaks = `, which can manually set the number of bars in the histogram.

```{r}

hist(dataset_sbp$SBP,
     breaks = 5)

```

```{r}

hist(dataset_sbp$SBP,
     breaks = 10)

```
We can change number of bars by setting `breaks = ` argument for `hist()` function.

## hist() - continued

By assigning `main = ` (title of histogram), `xlab = ` (x-axis label) and `ylab = ` (y-axis label), we can create a histogram with more detailed information.


```{r}

hist(dataset_sbp$SBP, 
     breaks = 10,
     main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
     xlab = "SBP (mmHg) of 1M Koreans",
     ylab = "Number of measurements"
     )

```

## hist() - percentage

By imputing `density` variable of histogram, we can create a histogram with y axis that is having percentage.


```{r}

h <- hist(dataset_sbp$SBP,
         breaks = 10, 
         plot = F)

h$density <- h$counts/sum(h$counts)*100

plot(h,
     freq = FALSE,
     main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
     xlab = "SBP (mmHg) of 1M Koreans",
     ylab = "Percentage of subject"
     )

```


# Boxplot

However, what should we do if we want to some summary statistic results as figures? Statisticians simply use boxplots. Boxplots can be generated by `boxplot()` command with information of what will be the x axis or colors.

## boxplot() - simple

We can directly put variable(vector) of our interest as `x`.

```{r}

boxplot(x = dataset_sbp$SBP,
        main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
        ylab = "SBP (mmHg)"
        )

```

## boxplot() - with x axis

Boxplot is useful in comparing data, by adding more information along x-axis.
To make more redundant boxplot, use `formula = y ~ x`. Here, `y` will be the variable of y-axis and x will be the x-axis.

**Remember x should be categorical (discrete) data.**

```{r}

boxplot(formula = SBP ~ SEX,
        data = dataset_sbp,
        main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
        ylab = "SBP (mmHg)",
        xlab = "Gender (male: 1, female: 2)"
        )

```

## boxplot() - x-axis label

we can also change x-axis texts with `names = ` argument.

```{r}

boxplot(SBP ~ SEX,
        data = dataset_sbp,
        main = "Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014",
        ylab = "SBP (mmHg)",
        xlab = "Gender",
        names = c("Male",
                  "Female")
        )

```

# Question

It seems like the histogram of SBP is somewhat having multiple peaks in the data. Can you tell why?

```{r}

hist(dataset_sbp$SBP,
     breaks = 500, 
     main = "Histogram of Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014 with 500 breaks",
     xlab = "SBP (mmHg) of 1M Koreans",
     ylab = "Number of measurements"
     )

```

# Common distributions

## Symmetric & bell-shaped

```{r}

hist(dataset_sbp$SBP,
     breaks = 10, 
     main = "Histogram of Systolic Blood Pressure (SBP)\nof 1M Koreans in 2013-2014 with 500 breaks",
     xlab = "SBP (mmHg) of 1M Koreans",
     ylab = "Number of measurements"
     )

```

## Right (positively) skewed

```{r}

hist(dataset_sbp$FBS,
     breaks = 10, 
     main = "Fasting Blood Sugar (FBS) levels\nof 1M Koreans in 2013-2014",
     xlab = "FBS (mg/L) of 1M Koreans",
     ylab = "Number of measurements"
     )

```
## Left (negatively skewed)


```{r}

hist(100-dataset_sbp$FBS,
     breaks = 10, 
     main = "100 - FBS",
     xlab = "100 - FBS",
     ylab = "Number of measurements"
     )

```

## Uniform distributions

```{r}
hist(dataset_sbp$BTH_G,
     breaks = 30, 
     main = "Age of cohort",
     xlab = "Age (years old)",
     ylab = "Number of measurements"
     )

```



# Closing remarks (swirl)

**How to learn basic R (optional)**

Install and run `swirl()`.

swirl teaches you R programming and data science interactively, at your own pace, and right in the R console!


```{r}
install.packages("swirl")
library(swirl)
```

Don't go too further,,, it will do almost the half of my job, teaching (bio)stats.

# Bibliography

```{r warning=FALSE, message=FALSE, echo=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
        str_libraries |>
                pacman::p_citation() |>
                print(bibtex = FALSE) |>
                capture.output() %>%
                .[-1:-3] %>% .[. != ""] |>
                stringr::str_squish() |>
                stringr::str_replace("_", "") |>
                cat()
        cat("\n")
}
#===============================================================================
```